Theories of deliberated choice

### 2024-04-30 simple.tex
rm claim about types

A claim about types of individuals is a set $C^\text{types} \wsub P^{\allhists}$. Given such a set, define the associated claim $C \wsub \allleadstos$ as $C = (C^\text{types})^I = \set{{\allleadstog} \in \allleadstos \suchthat \forall i \in I: {\ileadstog} \in C^\text{types}}$.
\begin{definition}
  Say that an attitude $g \in P^{\allhists}$ is observably included in a set of attitudes $S \wsub P^{\allhists}$ iff $\forall H \in \finite(\incr(\allhists)): \exists f \in S \suchthat \restr{g}{H} = \restr{f}{H}$.

  Define the extension of a claim about types $C^\text{types} \wsub P^{\allhists}$ to its observably included attitudes as the the claim about types $R(C^\text{types}) = \set{g \in P^{\allhists} \suchthat \forall H \in \finite(\incr(\allhists)): \exists f \in C^\text{types} \suchthat \restr{g}{H} = \restr{f}{H}}$.
  Note that $C^\text{types} \wsub R(C^\text{types})$. 
  Also note that $S_1 \wsub S_2 \implies R(S_1) \wsub R(S_2)$.

  Define the closure under observable inclusion of a claim about types $C^\text{types}$ as $R^\infty(C^\text{types})$, equivalently, as the smallest set $C^\text{types}_\text{ind}$ such that $R(C^\text{types}_\text{ind}) \wsub C^\text{types}_\text{ind}$.

@@

\begin{lemma}
  \label{th:typesfalsifiable}
  A claim about types of individuals $C^\text{types} \wsub P^{\allhists}$ 
  satisfies $\forall g \in P^{\allhists} \setminus C^\text{types}: \exists H \in \finite(\incr(\allhists)) \suchthat \forall f \in C^\text{types}: \restr{g}{H} \neq \restr{f}{H}$,
  iff its associated claim is falsifiable or trivial.
\end{lemma}
\begin{proof}
  Pick any ${\allleadstog} \notin C$.
  By definition of $C$,
  $\exists i \in I \suchthat {{\ileadstog}} \notin C^\text{types}$.
  As $C$ holds, ${\ileadstog} \notin \set{{\ileadsto[j]} \suchthat j \in I}$ (\cref{th:typesholds}).
  Pick any $H$ such that $\forall {\ileadstog[j]} \in C^\text{types}: \exists \hist \in H \suchthat {\ileadstog}(\hist) \neq {\ileadstog[j]}(\hist)$.
  Define $\theta = \set{(\hist, i, {\ileadstog}(\hist)) \suchthat \hist \in H} \in \Theta \cap \powerset{\allleadstog}$.
  Then, $\theta \notin C^\wsub$, because $\forall {\ileadstog[j]} \in C^\text{types}: \theta \nsubseteq {\ileadstog[j]}$.

  Conversely, assume that for some ${\ileadstog} \in P^{\allhists} \setminus C^\text{types}$, no $H \in \finite(\incr(\allhists))$ satisfies the indicated condition.
  Define ${\allleadstog}$ as including ${\ileadstog}$ and for all other individuals, relations from $C^\text{types}$.
  Then, ${\allleadstog} \notin C$, but any observation will be in $C$, thus the claim is not falsifiable and not trivial.
\end{proof}
\begin{lemma}
  \label{th:typesholds}
  Consider a claim about types of individuals $C^\text{types} \wsub P^{\allhists}$. Its associated claim holds iff $\set{{\ileadsto[j]} \suchthat j \in I} \wsub C^\text{types}$.
\end{lemma}
\begin{lemma}
  \label{th:typesanonymous}
  Given any claim about types of individuals $C^\text{types} \wsub P^{\allhists}$, its associated claim is anonymous.
\end{lemma}
The following result follows from \cref{th:typesfalsifiable,th:typesholds,th:typesanonymous}.
\begin{theorem}
  \label{th:suffFals}
  Assume that $\forall {\ileadstog} \in P^{\allhists} \setminus \set{{\ileadsto[k]} \suchthat k \in I}: \exists H \in \finite(\incr(\allhists)) \suchthat \forall j \in I: \exists \hist \in H \suchthat {\ileadstog}(\hist) \neq {\ileadsto[j]}(\hist)$.
  Then, the claim $C$ associated to the claim about types of individuals $C^\text{types} = \set{{\ileadsto[j]} \suchthat j \in I}$ holds, is anonymous, and is falsifiable or trivial.
\end{theorem}
\begin{proof}
  For another proof, note that $C^\wsub \cap \Theta = \set{\theta \in \Theta \suchthat \forall i \in I: \exists j \in I \suchthat \theta \cap (\allhists × \set{i} × P) \wsub {\ileadsto[j]}}$.
  Pick any ${\allleadstog} \in E(C^\wsub \cap \Theta)$ and let us show that ${\allleadstog} \in C$.
  Note that ${\allleadstog} \notin C \iff \exists i \in I \suchthat {\ileadstog} \in P^{\allhists} \setminus \set{{\ileadsto[k]} \suchthat k \in I}$, in which case the consequent in the theorem holds.
  We need to show that $\nexists i \in I, H \in \finite(\incr(\allhists)) \suchthat \forall j \in I: \exists \hist \in H \suchthat {\ileadstog}(\hist) \neq {\ileadsto[j]}(\hist)$.
  In other words, given any $i \in I$, we need to show that $\forall H \in \finite(\incr(\allhists)): \exists j \in I \suchthat \forall \hist \in H: {\ileadstog}(\hist) = {\ileadsto[j]}(\hist)$.
\end{proof}

The following result follows from \cref{th:typesfalsifiable}.
\begin{theorem}
  If a claim about types of individuals $C^\text{types} \wsub P^{\allhists}$ is not closed under observable inclusion, then it is not falsifiable and not trivial.
\end{theorem}
\begin{theorem}
  Given any situation, if the claim related to the claim about types of individuals $R^\infty({\allleadsto}(I))$ is trivial, then there exists no claim related to a claim about types of individuals that holds and is falsifiable.
\end{theorem}
\begin{proof}
  Consider a claim related to a claim about types of individuals $S$ that holds and that is falsifiable or trivial.
  For it to hold, we need ${\allleadsto}(I) \wsub S$.
  From \cref{th:typesfalsifiable}, for it to be falsifiable or trivial, we need that $\forall g \in P^{\allhists} \setminus S: \exists H \in \finite(\incr(\allhists)) \suchthat \forall f \in S: \restr{g}{H} \neq \restr{f}{H}$. In other words, all $g$ must be observably excludable from $S$. In other words, $S$ must be closed under observable inclusion. Thus, we need $R^\infty(S) \wsub S$.
  Also, because $S_1 \wsub S_2 \implies R(S_1) \wsub R(S_2)$, we need
  $R^\infty({\allleadsto}(I)) \wsub R^\infty(S) \wsub S$.
  It follows that $S$ is trivial.
\end{proof}
\begin{theorem}
  Given any situation, the claim about types of individuals $R^\infty({\allleadsto}(I))$ holds, is anonymous, and is falsifiable or trivial.
\end{theorem}

>>>>

Given a set of attitudes $S \wsub P^{\allhists}$, the claim associated to it designates the claim $S^I = \set{{\allleadstog} \in \allleadstos \suchthat \forall i \in I: {\ileadstog} \in S}$.
Say that a claim $C$ is a claim about types iff $\exists S \wsub P^{\allhists} \suchthat C = S^I$.
\begin{definition}
  Say that an attitude $g \in P^{\allhists}$ is observably included in a set of attitudes $S \wsub P^{\allhists}$ iff $\forall H \in \finite(\incr(\allhists)): \exists f \in S \suchthat \restr{g}{H} = \restr{f}{H}$.

  Define the extension of a set of attitudes $S \wsub P^{\allhists}$ to its observably included attitudes as the set of attitudes $R(S) = \set{g \in P^{\allhists} \suchthat \forall H \in \finite(\incr(\allhists)): \exists f \in S \suchthat \restr{g}{H} = \restr{f}{H}}$.
  Note that $S \wsub R(S)$. 
  Also note that $S_1 \wsub S_2 \implies R(S_1) \wsub R(S_2)$.

  Define the closure under observable inclusion of a claim about types $S$ as $R^\infty(S)$, equivalently, as the smallest set $S_\text{ind}$ such that $R(S_\text{ind}) \wsub S_\text{ind}$.

@@

\begin{lemma}
  \label{th:typesfalsifiable}
  A set of attitudes $S \wsub P^{\allhists}$ 
  satisfies $\forall g \in P^{\allhists} \setminus S: \exists H \in \finite(\incr(\allhists)) \suchthat \forall f \in S: \restr{g}{H} \neq \restr{f}{H}$,
  iff its associated claim is falsifiable or trivial.
\end{lemma}
\begin{proof}
  Pick any ${\allleadstog} \notin C$.
  By definition of $C$,
  $\exists i \in I \suchthat {{\ileadstog}} \notin S$.
  As $C$ holds, ${\ileadstog} \notin \set{{\ileadsto[j]} \suchthat j \in I}$ (\cref{th:typesholds}).
  Pick any $H$ such that $\forall {\ileadstog[j]} \in S: \exists \hist \in H \suchthat {\ileadstog}(\hist) \neq {\ileadstog[j]}(\hist)$.
  Define $\theta = \set{(\hist, i, {\ileadstog}(\hist)) \suchthat \hist \in H} \in \Theta \cap \powerset{\allleadstog}$.
  Then, $\theta \notin C^\wsub$, because $\forall {\ileadstog[j]} \in S: \theta \nsubseteq {\ileadstog[j]}$.

  Conversely, assume that for some ${\ileadstog} \in P^{\allhists} \setminus S$, no $H \in \finite(\incr(\allhists))$ satisfies the indicated condition.
  Define ${\allleadstog}$ as including ${\ileadstog}$ and for all other individuals, relations from $S$.
  Then, ${\allleadstog} \notin C$, but any observation will be in $C$, thus the claim is not falsifiable and not trivial.
\end{proof}
\begin{lemma}
  \label{th:typesholds}
  Consider a set of attitudes $S \wsub P^{\allhists}$. Its associated claim holds iff $\set{{\ileadsto[j]} \suchthat j \in I} \wsub S$.
\end{lemma}
\begin{lemma}
  \label{th:typesanonymous}
  Given any set of attitudes $S \wsub P^{\allhists}$, its associated claim is anonymous.
\end{lemma}
The following result follows from \cref{th:typesfalsifiable,th:typesholds,th:typesanonymous}.
\begin{theorem}
  \label{th:suffFals}
  Assume that $\forall {\ileadstog} \in P^{\allhists} \setminus \set{{\ileadsto[k]} \suchthat k \in I}: \exists H \in \finite(\incr(\allhists)) \suchthat \forall j \in I: \exists \hist \in H \suchthat {\ileadstog}(\hist) \neq {\ileadsto[j]}(\hist)$.
  Then, the claim $C$ associated to the set of attitudes $S = \set{{\ileadsto[j]} \suchthat j \in I}$ holds, is anonymous, and is falsifiable or trivial.
\end{theorem}
\begin{proof}
  For another proof, note that $C^\wsub \cap \Theta = \set{\theta \in \Theta \suchthat \forall i \in I: \exists j \in I \suchthat \theta \cap (\allhists × \set{i} × P) \wsub {\ileadsto[j]}}$.
  Pick any ${\allleadstog} \in E(C^\wsub \cap \Theta)$ and let us show that ${\allleadstog} \in C$.
  Note that ${\allleadstog} \notin C \iff \exists i \in I \suchthat {\ileadstog} \in P^{\allhists} \setminus \set{{\ileadsto[k]} \suchthat k \in I}$, in which case the consequent in the theorem holds.
  We need to show that $\nexists i \in I, H \in \finite(\incr(\allhists)) \suchthat \forall j \in I: \exists \hist \in H \suchthat {\ileadstog}(\hist) \neq {\ileadsto[j]}(\hist)$.
  In other words, given any $i \in I$, we need to show that $\forall H \in \finite(\incr(\allhists)): \exists j \in I \suchthat \forall \hist \in H: {\ileadstog}(\hist) = {\ileadsto[j]}(\hist)$.
\end{proof}

The following result follows from \cref{th:typesfalsifiable}.
\begin{theorem}
  If a set of attitudes $S \wsub P^{\allhists}$ is not closed under observable inclusion, then it is not falsifiable and not trivial.
\end{theorem}
\begin{theorem}
  Given any situation, if the claim associated to the set of attitudes $R^\infty({\allleadsto}(I))$ is trivial, then there exists no claim associated to a set of attitudes that holds and is falsifiable.
\end{theorem}
\begin{proof}
  Consider a claim related to a set of attitudes $S$ that holds and that is falsifiable or trivial.
  For it to hold, we need ${\allleadsto}(I) \wsub S$.
  From \cref{th:typesfalsifiable}, for it to be falsifiable or trivial, we need that $\forall g \in P^{\allhists} \setminus S: \exists H \in \finite(\incr(\allhists)) \suchthat \forall f \in S: \restr{g}{H} \neq \restr{f}{H}$. In other words, all $g$ must be observably excludable from $S$. In other words, $S$ must be closed under observable inclusion. Thus, we need $R^\infty(S) \wsub S$.
  Also, because $S_1 \wsub S_2 \implies R(S_1) \wsub R(S_2)$, we need
  $R^\infty({\allleadsto}(I)) \wsub R^\infty(S) \wsub S$.
  It follows that $S$ is trivial.
\end{proof}
\begin{theorem}
  Given any situation, the claim associated to the set of attitudes $R^\infty({\allleadsto}(I))$ holds, is anonymous, and is falsifiable or trivial.
\end{theorem}
